FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-06-02 paper.tex
cite

The IEEE 754 specification for floating-point numbers \cite{IEEEStandardBinary1985} governs how NaNs are handled as arguments to arithmetic functions.
Most of the time, when any or all of the arguments are NaN, the result is also a NaN.
This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.

@@

IEEE 754 \cite{IEEEStandardBinary1985} lists the following operations (among a few others) where NaNs result:

@@

For our first case study, we looked at the \texttt{ShallowWaters} library,\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021} which runs a simulates the flow of water over a seabed.
This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
That said, it works for our purposes as a proving ground for FloatTracker.
We dialed the CFL parameter up to high values (as high as 10) and we started seeing NaNs in the result.
Note that the program did not crash: instead, it began producing bad output.

>>>>

The IEEE 754 specification for floating-point numbers governs how NaNs are handled as arguments to arithmetic functions~\cite{IEEEStandardBinary1985}.
Most of the time, when any or all of the arguments are NaN, the result is also a NaN.
This is good and desirable, because it means that an arithmetic error---even if buried deep in an expression---will still be observable, allowing a user to find and debug it.

@@

IEEE 754~\cite{IEEEStandardBinary1985} lists the following operations (among a few others) where NaNs result:

@@

For our first case study, we looked at the \texttt{ShallowWaters} library~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021}, which runs a simulates the flow of water over a seabed.
This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.
That said, it works for our purposes as a proving ground for FloatTracker.
We dialed the CFL parameter up to high values (as high as 10) and we started seeing NaNs in the result.
Note that the program did not crash: instead, it began producing bad output.
